---
title: "HMM"
author: "Katharina Fijan"
date: "4/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE)
```


## Load Data & Packages
```{r}
library(tidyverse)
library(HMM)

data <- read.table('./data/TPP_proteasome_correlations.csv', sep=',', header = TRUE)
summary(data)

tpca_data <- read.table('~/Downloads/Proteasome_TPCA_Result.csv', sep=',', header =TRUE)
summary (tpca_data)
```

## Extract Data 
```{r}
psma1_df <- data %>% filter(bait == 'PSMA1', t!= 'asynch.median')
psma1_df <- psma1_df[order(psma1_df$prey),] ## each 5 entries is a new prey when ordered

psma2_df <- data %>% filter(bait == 'PSMA2', t!= 'asynch.median')
psma2_df <- psma2_df[order(psma2_df$prey),] ## each 5 entries is a new prey when ordered


##tpca data
psma1_df_tpca <- tpca_data %>% filter(bait == 'PSMA1', stage != 'asynch')
psma1_df_tpca <- psma1_df_tpca[order(psma1_df_tpca$prey),] 

# multi_obs <- paste(as.character(unlist(round(complex1_df$corr, 1))), as.character(unlist(round(complex2_df$corr, 1))), as.character(unlist(round(complex3_df$corr, 1))), as.character(unlist(round(complex4_df$corr, 1))), as.character(unlist(round(complex5_df$corr, 1))))

```


## Initialize HMM Model - Default Values
```{r}
#usage/defaults: initHMM(States, Symbols, startProbs=NULL, transProbs=NULL, emissionProbs=NULL)
states <- c("interaction", "no_interaction") #hidden states
symbols <- as.character(seq(from = -1.0, to = 1.0, by = 0.1)) #observed states - needs to be character dtype or error is thrown
pi <- NULL
T_ <- NULL
E_ <- NULL
hmm <- initHMM(states, symbols) #initialized model
print(hmm)
```


## Learn HMM - BaumWelch Algorithm
```{r}
#usage/defaults: baumWelch(hmm, observation, maxIterations=100, delta=1E-9, pseudoCount=0)

observed_data <- as.character(unlist(round(complex1_df$corr, 1))) #list of observed correlation values 
max_iterations <- 10000
delta=1E-9

learned_hmm <- baumWelch(hmm, observed_data, max_iterations, delta)
print(learned_hmm$hmm)
```


## Initialize HMM Model - Define Parameters
```{r}
#usage/defaults: initHMM(States, Symbols, startProbs=NULL, transProbs=NULL, emissionProbs=NULL)
states <- c("interaction", "no_interaction") #hidden states
symbols <- as.character(seq(from = 0.0, to = 1.0, by = 0.1)) #"correlation" values
pi <- c(0.5, 0.5)
T_ <- matrix(c(0.5, 0.5, 0.5, 0.5), nrow=2)
E_ <- learned_hmm$hmm$emissionProbs
hmm2<- initHMM(states, symbols, pi, T_, E_) #initialized model
print(hmm2)
```

## Learn HMM using Defined Params - BaumWelch Algorithm
```{r}
#usage/defaults: baumWelch(hmm, observation, maxIterations=100, delta=1E-9, pseudoCount=0)

observed_data <- as.character(unlist(abs(round(complex2_df$corr, 1)))) #list of observed correlation values 
max_iterations <- 10000
delta=1E-20

learned_hmm2 <- baumWelch(hmm2, observed_data)
print(learned_hmm2$hmm)
```



## Try aphid Package
```{r}
library(aphid)

#train(x, y, method = "Viterbi", seqweights = NULL, wfactor = 1, maxiter = 100, deltaLL = 1e-07,
 # logspace = "autodetect", quiet = FALSE, modelend = FALSE, pseudocounts = "Laplace", ...)

states <- c("begin", "interaction", "no_interaction")

### Define the transition probability matrix
T_ <- matrix(c(0, 0, 0, 0.5, 0.75, 0.75, 0.5, 0.25, 0.25), nrow = 3)
dimnames(T_) <- list(from = states, to = states) #names columns and rows for matrix

### Define the emission probability matrix
#Get the emission matrix for -1 to 1:
##Observations from -1 to 1
symbols = seq(from = -1.0, to = 1.0, by = 0.1)
##set an increasing probability from  0 to 1 spread out across the 20 observations
probs = seq(from = 0.0, to = 1.0, by = 0.05)
#Normalize so the rows will sum to 1 for each state
probs_increase_adjusted = probs * (1/sum(probs))
##get the reverse for the "no interaction" state
probs_decrease_adjusted = rev(probs_increase_adjusted)
##Set emission matrix
E_ <- matrix(c(probs_increase_adjusted, probs_decrease_adjusted), 2, byrow=TRUE)
#E_ <- matrix(c(rep(1/21, 21), rep(1/21, 21)), nrow = 2, byrow = TRUE)
dimnames(E_) <- list(states = states[-1], residues = symbols) #names columns and rows for matrix 

#creates hmm structure
x2 <- structure(list(A = T_, E = E_), class = "HMM")

#plots image of HMM
plot.HMM(x2, main ="Starting HMM Model for Protein Interactions")

observed_data <- c(as.character(round(psma1_df$corr, 1)))

#write function to loop through data and retrain model

#observed_data <- c(as.character(round(complex1_df$corr, 1)))
#dimnames(observed_data) <- list(states[1], states[1], states[1], states[0], states[0], states[1])

#trains hmm -- use 1 set of observations not multiple observations
hmm_aphid <- train(x2, observed_data, method="BaumWelch", maxiter=500, pseudocounts="Laplace")
hmm_aphid
hmm_aphid$A
hmm_aphid$E


plot.HMM(hmm_aphid, main="HMM Model After Training")
```

## Aphid & tpca Data
```{r}
library(aphid)

#train(x, y, method = "Viterbi", seqweights = NULL, wfactor = 1, maxiter = 100, deltaLL = 1e-07,
 # logspace = "autodetect", quiet = FALSE, modelend = FALSE, pseudocounts = "Laplace", ...)

states <- c("begin", "interaction", "no_interaction")

### Define the transition probability matrix
T_2 <- matrix(c(0, 0, 0, 0.5, 0.75, 0.75, 0.5, 0.25, 0.25), nrow = 3)
dimnames(T_2) <- list(from = states, to = states) #names columns and rows for matrix

### Define the emission probability matrix
#Get the emission matrix for -1 to 1:
##Observations from 0 to 1
symbols = seq(from = 0.0, to = 1.0, by = 0.1)
##set an increasing probability from  0 to 1 spread out across the 20 observations
# probs = seq(from = 0.0, to = 1.0, by = 0.1)
# #Normalize so the rows will sum to 1 for each state
# probs_increase_adjusted = probs * (1/sum(probs))
# ##get the reverse for the "no interaction" state
# probs_decrease_adjusted = rev(probs_increase_adjusted)
# ##Set emission matrix
#E_2 <- matrix(c(probs_increase_adjusted, probs_decrease_adjusted), 2, byrow=TRUE)

E_2 <- matrix(c(0, 1.0, 0.1, 0.9, 0.2, 0.8, 0.3, 0.7, 0.4, 0.6, 0.5, 0.5, 0.6, 0.4, 0.7, 0.3, 0.8, 0.2, 0.9, 0.1, 1.0, 0.0), nrow = 2) 

#E_ <- matrix(c(rep(1/21, 21), rep(1/21, 21)), nrow = 2, byrow = TRUE)
dimnames(E_2) <- list(states = states[2:3], residues = symbols) #names columns and rows for matrix 

#creates hmm structure
x_tpca <- structure(list(A = T_2, E = E_2), class = "HMM")

#plots image of HMM
plot.HMM(x2, main ="Starting HMM Model for Protein Interactions")

observed_data <- c(as.character(round(psma1_df_tpca$distance, 1)))
observed_df <- data.frame(actual = c(psma1_df_tpca$distance))
observed_df$binned <- ifelse(observed_df$actual > 1, '1.0', observed_df$actual) 

tpca1_train <- c(as.character(round(as.numeric(observed_df$binned, 1))))


#write function to loop through data and retrain model

#observed_data <- c(as.character(round(complex1_df$corr, 1)))
#dimnames(observed_data) <- list(states[1], states[1], states[1], states[0], states[0], states[1])

#trains hmm -- use 1 set of observations not multiple observations
hmm_aphid <- train(x_tpca, tpca1_train, method="BaumWelch", maxiter=500, pseudocounts="Laplace")
hmm_aphid
hmm_aphid$A
hmm_aphid$E

plot.HMM(hmm_aphid, main="HMM Model After Training")



hmm_tpcaE <- data.frame(t(hmm_aphid$E))

ggplot(data=hmm_tpcaE) + 
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=interaction), fill='coral1',color='coral4') + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=interaction), color='coral') +
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=no_interaction), fill='cornflowerblue', color='darkslateblue') + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=no_interaction), color='blue') +
  ggtitle("Learned Emissions Matrix") + 
  xlab("Correlation Values") +
  ylab("Probabilities") + 
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=data.frame(t(E_2))) + 
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=interaction), fill='coral1',color='coral4') + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=interaction), color='coral') +
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=no_interaction), fill='cornflowerblue', color='darkslateblue') + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=no_interaction), color='blue') +
  ggtitle("Initialized Emissions Matrix") + 
  xlab("Correlation Values") +
  ylab("Probabilities") + 
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

```


### Try learning HMM using pvalues from exosome data
```{r}
#train(x, y, method = "Viterbi", seqweights = NULL, wfactor = 1, maxiter = 100, deltaLL = 1e-07,
 # logspace = "autodetect", quiet = FALSE, modelend = FALSE, pseudocounts = "Laplace", ...)

states <- c("begin", "interaction", "no_interaction")

### Define the transition probability matrix
T_2 <- matrix(c(0, 0, 0, 0.5, 0.75, 0.75, 0.5, 0.25, 0.25), nrow = 3)
dimnames(T_2) <- list(from = states, to = states) #names columns and rows for matrix

### Define the emission probability matrix
#Get the emission matrix for -1 to 1:
##Observations from 0 to 1

E_2 <- matrix(c(0, 1.0, 0.1, 0.9, 0.2, 0.8, 0.3, 0.7, 0.4, 0.6, 0.5, 0.5, 0.6, 0.4, 0.7, 0.3, 0.8, 0.2, 0.9, 0.1, 1.0, 0.0), nrow = 2) 

#E_ <- matrix(c(rep(1/21, 21), rep(1/21, 21)), nrow = 2, byrow = TRUE)
dimnames(E_2) <- list(states = c(states[3], states[2]), residues = symbols) #names columns and rows for matrix 

#creates hmm structure
x_tpca <- structure(list(A = T_2, E = E_2), class = "HMM")

#plots image of HMM
plot.HMM(x2, main ="Starting HMM Model for Protein Interactions")

pval <- tpca_pvalue_data %>% filter(bait == 'EXOSC1', stage!= 'asynch')
pval <- pval[order(pval$prey),]
observed_data <- c(as.character(round(tpca_pvalue_data$pvalue, 1)))
observed_df <- data.frame(actual = c(tpca_pvalue_data$pvalue))
#observed_df$binned <- ifelse(observed_df$actual > 1, '1.0', observed_df$actual) 

tpca1_train <- c(as.character(round(as.numeric(observed_df$actual, 1))))

#write function to loop through data and retrain model

#observed_data <- c(as.character(round(complex1_df$corr, 1)))
#dimnames(observed_data) <- list(states[1], states[1], states[1], states[0], states[0], states[1])

#trains hmm -- use 1 set of observations not multiple observations
hmm_aphid <- train(x_tpca, tpca1_train, method="BaumWelch", maxiter=500, pseudocounts="Laplace")
hmm_aphid
hmm_aphid$A
hmm_aphid$E

plot.HMM(hmm_aphid, main="HMM Model After Training")


hmm_tpcaE <- data.frame(t(hmm_aphid$E))

ggplot(data=hmm_tpcaE) + 
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=interaction), fill='coral1',color='coral4') + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=interaction), color='coral') +
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=no_interaction), fill='cornflowerblue', color='darkslateblue') + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=no_interaction), color='blue') +
  ggtitle("Learned Emissions Matrix") + 
  xlab("Correlation Values") +
  ylab("Probabilities") + 
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=data.frame(t(E_2))) + 
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=interaction), fill='coral1',color='coral4') + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=interaction), color='coral') +
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=no_interaction), fill='cornflowerblue', color='darkslateblue') + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.1), y=no_interaction), color='blue') +
  ggtitle("Initialized Emissions Matrix") + 
  xlab("Correlation Values") +
  ylab("Probabilities") + 
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```




```{r}
x <- seq(0, 0.5, by=0.01)
noncor <- dnorm(x, mean= mean(x), sd(x) )
x2 <- seq(0.3, 1.0, by=0.01)
cor <- dnorm(x2, mean=mean(x2), sd(x2))

ggplot() +
  geom_line(aes(x, noncor), fill='cornflowerblue', color='darkslateblue') +
  geom_line(aes(x2, cor),color='coral3') +
  geom_vline(xintercept= .44, col = 'red') +
  xlab("Correlation") +
  ylab("") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Gaussian Mixture Model (Illustrative Example)")

```


```{r}
#try deriveHMM
#add interaction labels
psma1_df_tpca$pro_interaction <- ifelse(psma1_df_tpca$distance > 0.3, 'interaction', 'no_interaction') 
psma1_df_tpca$distance <- ifelse(psma1_df_tpca$distance >1.0, 1.0, psma1_df_tpca$distance)

named_vector1 <- c(as.character(round(psma1_df_tpca$distance, 1)))
names(named_vector1) <- c(as.character(psma1_df_tpca$pro_interaction))

derive_tpca <- deriveHMM(list(named_vector1), logspace = FALSE)

derive_tpca$A
derive_tpca$E
plot.HMM(derive_tpca)

derive_E <- data.frame(t(derive_tpca$E))

ggplot(data=derive_E) + 
  geom_point(aes(x = seq(from = 0.1, to = 1.0, by = 0.1), y=interaction), fill='coral1',color='coral4') + 
  geom_line(aes(x = seq(from = 0.1, to = 1.0, by = 0.1), y=interaction), color='coral') +
  geom_point(aes(x = seq(from = 0.1, to = 1.0, by = 0.1), y=no_interaction), fill='cornflowerblue', color='darkslateblue') + 
  geom_line(aes(x = seq(from = 0.1, to = 1.0, by = 0.1), y=no_interaction), color='blue') +
  ggtitle("Learned Emissions Matrix") + 
  xlab("Correlation Values") +
  ylab("Probabilities") + 
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

```


```{r}
#try deriveHMM
#add interaction labels
psma1_df$pro_interaction <- ifelse(psma1_df$corr > 0.5, 'interaction', 'no_interaction')

named_vector <- c(as.character(round(psma1_df$corr, 1)))
names(named_vector) <- c(as.character(psma1_df$pro_interaction))

deriveh <- deriveHMM(list(named_vector), logspace = FALSE)

deriveh$A
deriveh$E
plot.HMM(deriveh)
```




```{r}
hmm_dfT <- data.frame(hmm_aphid$A) 
hmm_dfE <- data.frame(t(hmm_aphid$E))

ggplot(data=hmm_dfE) + 
  geom_col(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=interaction), fill='coral1',color='coral4', alpha = 0.3) +
  geom_col(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=no_interaction), fill='cornflowerblue', color='darkslateblue', alpha = 0.2) +
  ggtitle("Learned Emissions Matrix") + 
  xlab("Correlation Values") +
  ylab("Probabilities") + 
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))



ggplot(data=data.frame(t(E_))) +
  geom_col(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=interaction), fill='coral1',color='coral4', alpha = 0.3) +
  geom_col(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=no_interaction), fill='cornflowerblue', color='darkslateblue', alpha = 0.2) +
  ggtitle("Initialized Emissions Matrix") + 
  xlab("Correlation Values") +
  ylab("Probabilities") + 
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data=hmm_dfE) + 
  geom_point(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=interaction), fill='coral1',color='coral4') + 
  geom_line(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=interaction), color='coral') +
  geom_point(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=no_interaction), fill='cornflowerblue', color='darkslateblue') + 
  geom_line(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=no_interaction), color='blue') +
  ggtitle("Learned Emissions Matrix") + 
  xlab("Correlation Values") +
  ylab("Probabilities") + 
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=data.frame(t(E_))) + 
  geom_point(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=interaction), fill='coral1',color='coral4') + 
  geom_line(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=interaction), color='coral') +
  geom_point(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=no_interaction), fill='cornflowerblue', color='darkslateblue') + 
  geom_line(aes(x = seq(from = -1.0, to = 1.0, by = 0.1), y=no_interaction), color='blue') +
  ggtitle("Initialized Emissions Matrix") + 
  xlab("Correlation Values") +
  ylab("Probabilities") + 
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))


```

# Viterbi to determine protein states
```{r}

test_viterbi <- c(as.character(round(psma2_df$corr, 1)))

vit1 <- Viterbi(hmm_aphid, test_viterbi)
predicted <- c("interaction", "no_interaction")[vit1$path +1]
predicted

# protein.post <- posterior(hmm_aphid, observed_data)
# plot(1:300, seq(0, 1, length.out = 300), type = "n", xlab = "Protein's",
#      ylab = "Posterior probability of Interaction or No Interaction")
# starts <- which(c("L", actual) == "F" & c(actual, "F") == "L")
# ends <- which(c("F", actual) == "L" & c(actual, "L") == "F") - 1
# for(i in 1:6) rect(starts[i], 0, ends[i], 1, col = "grey", border = NA)
# lines(1:300, casino.post[1, ])
```


## train on psma2 to see if different parameters found
```{r}

x_base <- structure(list(A = T_, E = E_), class = "HMM")

hmm_aphid2 <- train(x_base, test_viterbi, method="BaumWelch", maxiter=500, pseudocounts="Laplace")
hmm_aphid2
hmm_aphid2$A
hmm_aphid2$E

vit2 <- Viterbi(hmm_aphid2, test_viterbi)
predicted2 <- c("interaction", "no_interaction")[vit2$path +1]
predicted2
```

## loop through different protein interactions - psma1
```{r}
observed_data <- c(as.character(round(psma1_df$corr, 1)))

#write function to loop through data and retrain model
hmm_looped <- structure(list(A = T_, E = E_), class = "HMM") #base model
index1 <- 1
for (pairing in range((length(observed_data)/5) + 1)) {
  pair_train <- c(as.character(round(psma1_df$corr, 1)[index1:(index1+5)]))
  hmm_looped <- train(hmm_looped, observed_data, method="BaumWelch", maxiter=500, pseudocounts="Laplace")
  index1 <- index1 + 5
}
hmm_looped$A
hmm_looped$E
plot.HMM(hmm_looped, main="HMM Model After Training")
```



## Playing around with aphid example data/code -- casino 

```{r}
data(casino) ##data used in examples

## the dishonest casino example from Durbin et al (1998)
states <- c("Begin", "Fair", "Loaded")
residues = paste(1:6)
A <- matrix(c(0, 0, 0, 0.99, 0.95, 0.1, 0.01, 0.05, 0.9), nrow = 3)
dimnames(A) <- list(from = states, to = states)
E <- matrix(c(rep(1/6, 6), rep(1/10, 5), 1/2), nrow = 2, byrow = TRUE)
dimnames(E) <- list(states = states[-1], residues = residues)
x <- structure(list(A = A, E = E), class = "HMM")
plot(x, main = "Dishonest casino hidden Markov model")

##deriveHMM 
derive  <- deriveHMM(list(casino))
derive$A
derive$E
```

```{r}
## Baum Welch training for standard HMMs:
  ## The dishonest casino example from Durbin et al (1998) chapter 3.2
states <- c("Begin", "Fair", "Loaded")
residues <- paste(1:6)
### Define the transition probability matrix
A <- matrix(c(0, 0, 0, 0.99, 0.95, 0.1, 0.01, 0.05, 0.9), nrow = 3)
dimnames(A) <- list(from = states, to = states)
### Define the emission probability matrix
E <- matrix(c(rep(1/6, 6), rep(1/10, 5), 1/2), nrow = 2, byrow = TRUE)
dimnames(E) <- list(states = states[-1], residues = residues)
### Build and plot the HMM object
x <- structure(list(A = A, E = E), class = "HMM")
op <- par(no.readonly = TRUE)
par(mfrow = c(2, 1))
plot(x, main = "Dishonest casino HMM before training")
data(casino)
x <- train(x, list(casino), method = "BaumWelch", deltaLL = 0.001)
plot(x, main = "Dishonest casino HMM after training")
par(op)
```


## By Hand - try? 
```{r}
# #update H,L,M to 1,2,3 for function. 
# spring$park_obs <- ifelse(spring$parks < -10, 1,
#                           ifelse(spring$parks >= -10 & spring$parks < 40, 2, 3))
# 
# #BaumWelch
# BaumWelch = function(v, a, b, initial_distribution, n.iter = 100){
#   
#   for(i in 1:n.iter){
#     T = length(v)
#     M = nrow(a)
#     K=ncol(b)
#     alpha = forward(v, a, b, initial_distribution)
#     beta = backward(v, a, b)
#     xi = array(0, dim=c(M, M, T-1))
#     
#     for(t in 1:T-1){
#       denominator = ((alpha[t,] %*% a) * b[,v[t+1]]) %*% matrix(beta[t+1,]) 
#       for(s in 1:M){
#         numerator = alpha[t,s] * a[s,] * b[,v[t+1]] * beta[t+1,]
#         xi[s,,t]=numerator/as.vector(denominator)
#       }
#     }
#     
#     
#     xi.all.t = rowSums(xi, dims = 2)
#     a = xi.all.t/rowSums(xi.all.t)
#     
#     gamma = apply(xi, c(1, 3), sum)  
#     gamma = cbind(gamma, colSums(xi[, , T-1]))
#     for(l in 1:K){
#       b[, l] = rowSums(gamma[, which(v==l)])
#     }
#     b = b/rowSums(b)
#     
#   }
#   return(list(a = a, b = b, initial_distribution = initial_distribution))
# }
# 
# M=3; K=3
# A = matrix(1, M, M)
# A = A/rowSums(A)
# B = matrix(1:6, M, K)
# B = B/rowSums(B)
# initial_distribution = c(1/3,1/3, 1/3)
# 
# (myout = BaumWelch(spring$park_obs, A, B, initial_distribution, n.iter = 100))
```


