---
title: "HMM TPP Preprocessing"
output: html_notebook
---

**Loading Inputs**

#Library import
```{r}
library(readxl)
library(data.table)
library(dplyr)
library(Hmisc)
library(ggplot2)

#Rtpca specific
library(Rtpca)
library(eulerr)
```




#Search Annotation
Criteria fot limiting the genes that will be used to find interactions
```{r}
#Ribosome
large_ribosomal = "RPL"
small_ribosomal = "RPS"


#Proteasome
core_proteasome = c("PSMA", "PSMB")
regulatory_proteasome = c("PSMC", "PSMD")
alt_regulatory_proteasome = "PSME"
proteasome_assembly_chaperone = "PSMG"


#Exosome
core_exosome = c("EXOSC1", "EXOSC4", "EXOSC5", "EXOSC6", "EXOSC8", "EXOSC9", "EXOSC2", "EXOSC7")
reg_exosome = c("DIS3", "EXOSC10")

filter_criteria = c(core_exosome, reg_exosome)
```


**Pearson Analysis**

#Import and shape sample data
```{r}
sample_data = read_excel("data/TPP_sampleData.xlsx", sheet = "TableS1_2D-TPP")
View(sample_data)


sample_data = as.data.table(sample_data[,c("gene_name", "description", "Temperature", "earlyS.median", "lateS.median", "S_G2.median", "M.median", "G1.median", "asynch.median")])

sample_data = melt(sample_data, id.vars = c("gene_name", "description", "Temperature"), variable.name = "t", value.name = "median_fx")
```

#Filter sample data
Filters the sample data based on the previously determined filter criteria
```{r}
filter_data = function(data, filter, contain = TRUE) {
  
  target_names = c()
  
  if (contain == TRUE){
    
    for (i in filter) {
      x = data[grep(i, data$gene_name),]
      target_names = c(target_names, unique(x$gene_name))
    }
    
  } else {
    target_names = filter
  }
  return(target_names)
}

target_proteins = filter_data(sample_data, filter_criteria, contain = TRUE)
target_proteins
```


#Initialize DF
The following function initializes a correlation dataframe for downstream analysis by converting the data into pairwise interactions.
```{r}
#initialize df

Initialize_CorDF = function(targets, data) {
  
  for (i in 1:(length(targets)-1)){ #1:

    if (i == 1){
      corr_df = data.frame(bait = targets[i], prey = c(targets[(i+1):length(targets)]))
    } else {
      new_row = data.frame(bait = targets[i], prey = c(targets[(i+1):length(targets)]))
      corr_df = bind_rows(corr_df, new_row)
    }
    
  }
  
  time_points = data.frame(t = unique(data$t))
  corr_df = merge(corr_df, time_points)
  corr_df$corr = 0
  
  return(corr_df)
}


corr_df = Initialize_CorDF(target_proteins, sample_data)
corr_df

```


#Find all correlation values
The following function adds a pairwise correlation between the bait and the prey
```{r, warning=FALSE}
corr_table = corr_df

Append_Cor = function(corr_table, data) {
  
  for (i in 1:(nrow(corr_table)+1)) { #
    
    #check progress
    if (i %% 100 == 0) {
      print(paste0(i,"/",nrow(corr_table)))
    }
    
    #first filter by stage
    time_point = corr_table[i,"t"]
    
    #Filter the bait
    bait = corr_table[i,"bait"]
    corr_data_bait = filter(data, t == time_point) %>% filter(gene_name == bait)
    
    #Filter the prey
    prey = corr_table[i,"prey"]
    corr_data_prey = filter(data, t == time_point) %>% filter(gene_name == prey)
    
    #Find the correlation
    
    corr = rcorr(as.numeric(corr_data_bait$median_fx), as.numeric(corr_data_prey$median_fx))
    #print()
  
    #Add correlation into corrrelation table
    corr_table[i, "corr"] = corr$r["x","y"]
  }
  
  return(corr_table)
}

corr_table = Append_Cor(corr_df, sample_data)

```
```{r}
View(corr_table)
```


#Run all functions
```{r}
filter_criteria = c(core_proteasome, regulatory_proteasome)

target_proteins = filter_data(sample_data, filter_criteria, contain = TRUE)
target_proteins

corr_df = Initialize_CorDF(target_proteins, sample_data)
corr_df

corr_table = Append_Cor(corr_df, sample_data)
corr_table

write.csv(corr_table, file = "data/TPP_exosome_correlations.csv", quote = FALSE, row.names = FALSE, col.names = TRUE)
```




**Rtpca Analysis**

#Load data
```{r}
data("ori_et_al_complexes_df")
#ori_et_al_complexes_df

#supp_tab_becher_s4 <- read_xlsx("1-s2.0-S0092867418303854-mmc4.xlsx",sheet = "TableS4_TPP-TR")

becher = read.csv("data/becher_fullCycle_reformat.csv")
becher

#unique(becher$cell.cycle)
```

#Prepare data
```{r}
RTPCA_Preprocess = function(cell_cycle, becher_data) {
  
  #Reformat the dataframe
  reformat_df <- becher_data %>% 
    filter(cell.cycle == cell_cycle) %>% 
    dplyr::select(
        gene_name,
        replicates = found.in.reps,
        max_qupm = max.qupm,
        min_qupm = min.qupm,
        matches("mean\\.fc")) %>% 
    filter(min_qupm > 0, replicates > 1)
  
  #And we create as matrix, define its row names and supply and attribute vector specifying the temperatures represented by the various TMT channels:
  
  reformat_mat <- as.matrix(reformat_df %>% dplyr::select(dplyr::matches("mean\\.fc")))
  
  rownames(reformat_mat) <- reformat_df$gene_name
  
  #Annotate temperature
  temperature_anno <- 
      as.numeric(
          gsub("T", "", gsub("_.+", "", colnames(
              becher_data %>% 
                  dplyr::select(matches("mean\\.fc"))))))
  
  attributes(reformat_mat)$temperature <- temperature_anno
  
  return(reformat_mat)
  
}

earlyS_mat = RTPCA_Preprocess("earlyS", becher)
earlyS_mat
```


#Distance matrix functions
```{r}
Find_Distances = function(cell_cycle, preprocessed_data, complex_data, filter, match = FALSE) {
  
  #Filter for complex of interest
  #complex_proteins = filter(complex_data, id == "26S Proteasome")
  
  #Run Tpca analysis
  TPCA <- runTPCA(objList = list(preprocessed_data), complexAnno = complex_data)
  plotComplexRoc(TPCA, computeAUC = TRUE)
  
  #Extract distance matrix and filter columns
  if (match == TRUE) {
    target_distance_matrix_all = data.frame(TPCA@DistMat) %>% select(ends_with(filter))
    #Filter the rows as well
    target_distance_matrix = select(data.frame(t(target_distance_matrix_all)), ends_with(filter)) %>% as.matrix()
  }  else {
    target_distance_matrix_all = data.frame(TPCA@DistMat) %>% select(contains(filter))
    #Filter the rows as well
    target_distance_matrix = select(data.frame(t(target_distance_matrix_all)), contains(filter)) %>% as.matrix()
  }
  
  #Remove duplicates
  target_distance_matrix[lower.tri(target_distance_matrix)] = 0
  
  #Transform matrix into table
  interaction_table = melt(target_distance_matrix) %>% filter(value != 0)
  interaction_table$stage = cell_cycle
  colnames(interaction_table) = c("bait", "prey", "distance", "stage")
  
  
  #Add pvalue
  ##Select only null interactions and convert to table
  null_distance_matrix = select(data.frame(t(target_distance_matrix_all)), !contains(filter_criteria)) %>% as.matrix()
  null_table = reshape2::melt(null_distance_matrix) %>% filter(value != 0)
  ##Derive the pvalues and add to interactions table
  pvalues <- empPvals(stat = interaction_table$distance, stat0 = null_table$value)
  interaction_table$pvalue = (1 - pvalues)
  
  return(interaction_table)
}

distance_table = Find_Distances("earlyS", earlyS_mat, ori_et_al_complexes_df, filter_criteria, match = TRUE)

View(distance_table)
```


#Run all functions and combine into a single DF
```{r}
CorrAnalysis.TPCA = function(time_points, data, complex_data, filter, match = FALSE) {
  
  output_table = data.frame()
  
  for (t in time_points){
    #Preprocess data for Rtpca analysis
    preprocess_mat = RTPCA_Preprocess(t, data)
    
    #Run Rptca and find the distance matrices & reformat data
    distance_table = Find_Distances(t, preprocess_mat, complex_data, filter, match)
    
    #Combine matrices into single dataframe
    output_table = bind_rows(output_table, distance_table)
  }
  return(output_table)
}

cell_cycle = c("G1", "earlyS", "lateS", "S_G2", "M", "asynch")

TPCA_Result = CorrAnalysis.TPCA(cell_cycle, becher, ori_et_al_complexes_df, filter_criteria, match = TRUE)

View(TPCA_Result)

write.csv(TPCA_Result, "Exosome_TPCA_Result_pvalues.csv", row.names = FALSE, quote = FALSE)
```
