---
title: "Data_Preprocessing"
output: html_notebook
---

**Loading Inputs**

#Library import
```{r}
library(readxl)
library(data.table)
library(dplyr)
library(Hmisc)
```

#Import and shape sample data
```{r}
sample_data = read_excel("data/TPP_sampleData.xlsx", sheet = "TableS1_2D-TPP")

sample_data = as.data.table(sample_data[,c("gene_name", "description", "Temperature", "earlyS.median", "lateS.median", "S_G2.median", "M.median", "G1.median", "asynch.median")])

sample_data = melt(sample_data, id.vars = c("gene_name", "description", "Temperature"), variable.name = "t", value.name = "median_fx")
```


#Search Annotation
Criteria fot limiting the genes that will be used to find interactions
```{r}
#Ribosome
large_ribosomal = "RPL"
small_ribosomal = "RPS"


#Proteasome
core_proteasome = c("PSMA", "PSMB")
regulatory_proteasome = c("PSMC", "PSMD")
alt_regulatory_proteasome = "PSME"
proteasome_assembly_chaperone = "PSMG"


#Exosome
core_exosome = c("EXOSC1", "EXOSC4", "EXOSC5", "EXOSC6", "EXOSC8", "EXOSC9", "EXOSC2", "EXOSC7")
reg_exosome = c("DIS3", "EXOSC10")

```


**Analysis Functions**

#Filter sample data
Filters the sample data based on the previously determined filter criteria
```{r}
filter_data = function(data, filter, contain = TRUE) {
  
  target_names = c()
  
  if (contain == TRUE){
    
    for (i in filter) {
      x = data[grep(i, data$gene_name),]
      target_names = c(target_names, unique(x$gene_name))
    }
    
  } else {
    target_names = filter
  }
  return(target_names)
}

filter_criteria = c(core_exosome, reg_exosome)

target_proteins = filter_data(sample_data, filter_criteria, contain = TRUE)
target_proteins
```


#Initialize DF
The following function initializes a correlation dataframe for downstream analysis by converting the data into pairwise interactions.
```{r}
#initialize df

Initialize_CorDF = function(targets, data) {
  
  for (i in 1:(length(targets)-1)){ #1:

    if (i == 1){
      corr_df = data.frame(bait = targets[i], prey = c(targets[(i+1):length(targets)]))
    } else {
      new_row = data.frame(bait = targets[i], prey = c(targets[(i+1):length(targets)]))
      corr_df = bind_rows(corr_df, new_row)
    }
    
  }
  
  time_points = data.frame(t = unique(data$t))
  corr_df = merge(corr_df, time_points)
  corr_df$corr = 0
  
  return(corr_df)
}


corr_df = Initialize_CorDF(target_proteins, sample_data)
corr_df

```


#Find all correlation values
The following function adds a pairwise correlation between the bait and the prey
```{r, warning=FALSE}
corr_table = corr_df

Append_Cor = function(corr_table, data) {
  
  for (i in 1:nrow(corr_table)) { #
    
    #check progress
    if (i %% 100 == 0) {
      print(paste0(i,"/",nrow(corr_table)))
    }
    
    #first filter by stage
    time_point = corr_table[i,"t"]
    
    #Filter the bait
    bait = corr_table[i,"bait"]
    corr_data_bait = filter(data, t == time_point) %>% filter(gene_name == bait)
    
    #Filter the prey
    prey = corr_table[i,"prey"]
    corr_data_prey = filter(data, t == time_point) %>% filter(gene_name == prey)
    
    #Find the correlation
    
    corr = rcorr(as.numeric(corr_data_bait$median_fx), as.numeric(corr_data_prey$median_fx))
    #print()
  
    #Add correlation into corrrelation table
    corr_table[i, "corr"] = corr$r["x","y"]
  }
  
  return(corr_table)
}

corr_table = Append_Cor(corr_df, sample_data)

```


#Run all functions
```{r}
filter_criteria = c(core_proteasome, regulatory_proteasome)

target_proteins = filter_data(sample_data, filter_criteria, contain = TRUE)
target_proteins

corr_df = Initialize_CorDF(target_proteins, sample_data)
corr_df

corr_table = Append_Cor(corr_df, sample_data)
corr_table
```

#Save dataframe
```{r}
write.csv(corr_table, file = "data/TPP_exosome_correlations.csv", quote = FALSE, row.names = FALSE, col.names = TRUE)
```