---
title: "HMM Analysis Vignette"
output: html_notebook
---

**Section 1: Loading Data**

#Loading packages
For any packages not already installed, install via the 'install.packages()' function. 
Note: The Rtpca requires specific versioning and other packages, if prompted install and version accordingly. 
```{r}
library(tidyr)
library(dplyr)
library(aphid)
library(tidyverse)
library(HMM)
library(readxl)
library(data.table)
library(Hmisc)
library(ggplot2)
library(Rtpca) 
library(eulerr)
library(qvalue)
```

#Loading data
Data loaded is retrieved from Becher et al. (2018): https://data.mendeley.com/datasets/xrbmvv5srs/2. 
```{r}
#Protein TPP data
becher = read.csv("data/becher_fullCycle_reformat.csv")

#Protein Complex annotation imported with the Rtpca package
data("ori_et_al_complexes_df")

head(becher, n=6)
```


**Section 2: Preprocessing**

#Prepare data
Resturctures the primary becher data to fit the RTPCA analysis function for the Residual Sum of Squares funcion.
```{r}
RTPCA_Preprocess = function(cell_cycle, becher_data) {
  
  #Reformat the dataframe
  reformat_df <- becher_data %>% 
    filter(cell.cycle == cell_cycle) %>% 
    dplyr::select(
        gene_name,
        replicates = found.in.reps,
        max_qupm = max.qupm,
        min_qupm = min.qupm,
        matches("mean\\.fc")) %>% 
    filter(min_qupm > 0, replicates > 1)
  
  #Create as matrix, define its row names and supply and attribute vector specifying the temperatures represented by the various TMT channels:
  
  reformat_mat <- as.matrix(reformat_df %>% dplyr::select(dplyr::matches("mean\\.fc")))
  
  rownames(reformat_mat) <- reformat_df$gene_name
  
  #Annotate temperature
  temperature_anno <- 
      as.numeric(
          gsub("T", "", gsub("_.+", "", colnames(
              becher_data %>% 
                  dplyr::select(matches("mean\\.fc"))))))
  
  attributes(reformat_mat)$temperature <- temperature_anno
  
  return(reformat_mat)
  
}

##To test individual function run the code below on a single time point:
#RTPCA_Preprocess("earlyS", becher)
```

#Filter Criteria
Define the protein complex of interest by listing the part of or the entire protein name. 
```{r}
#Exosome
core_exosome = c("EXOSC1", "EXOSC4", "EXOSC5", "EXOSC6", "EXOSC8", "EXOSC9", "EXOSC2", "EXOSC7")
reg_exosome = c("DIS3", "EXOSC10")

filter_criteria = c(core_exosome, reg_exosome)
```

#Distance matrix functions
Find the Residual Sum of Squares between all the proteins and then filter by specific complex of interest.
```{r}
Find_Distances = function(cell_cycle, preprocessed_data, complex_data, filter, match = FALSE) {
  
  #Filter for complex of interest
  #complex_proteins = filter(complex_data, id == "26S Proteasome")
  
  #Run Tpca analysis
  TPCA <- runTPCA(objList = list(preprocessed_data), complexAnno = complex_data)
  plotComplexRoc(TPCA, computeAUC = TRUE)
  
  #Extract distance matrix and filter columns
  if (match == TRUE) {
    target_distance_matrix_all = data.frame(TPCA@DistMat) %>% select(ends_with(filter))
    #Filter the rows as well
    target_distance_matrix = select(data.frame(t(target_distance_matrix_all)), ends_with(filter)) %>% as.matrix()
  }  else {
    target_distance_matrix_all = data.frame(TPCA@DistMat) %>% select(contains(filter))
    #Filter the rows as well
    target_distance_matrix = select(data.frame(t(target_distance_matrix_all)), contains(filter)) %>% as.matrix()
  }
  
  #Remove duplicates
  target_distance_matrix[lower.tri(target_distance_matrix)] = 0
  
  #Transform matrix into table
  interaction_table = melt(target_distance_matrix) %>% filter(value != 0)
  interaction_table$stage = cell_cycle
  colnames(interaction_table) = c("bait", "prey", "distance", "stage")
  
  #Add pvalue
  ##Select only null interactions and convert to table
  null_distance_matrix = select(data.frame(t(target_distance_matrix_all)), !contains(filter_criteria)) %>% as.matrix()
  null_table = melt(null_distance_matrix) %>% filter(value != 0)
  ##Derive the pvalues and add to interactions table
  pvalues <- empPvals(stat = interaction_table$distance, stat0 = null_table$value)
  interaction_table$pvalue = (1 - pvalues)
  
  return(interaction_table)
}

##To test individual function run the code below on a single time point:
#Find_Distances("earlyS", earlyS_mat, ori_et_al_complexes_df, filter_criteria, match = TRUE)
```

#Run all functions and combine into a single DF
Run the entire preprocessing altogether for input into the HMM model
Note: Will take 5-10 minutes (using i7 4-core processor).
```{r}
CorrAnalysis.TPCA = function(time_points, data, complex_data, filter, match = FALSE) {
  
  output_table = data.frame()
  
  for (t in time_points){
    #Preprocess data for Rtpca analysis
    preprocess_mat = RTPCA_Preprocess(t, data)
    
    #Run Rptca and find the distance matrices & reformat data
    distance_table = Find_Distances(t, preprocess_mat, complex_data, filter, match)
    
    #Combine matrices into single dataframe
    output_table = bind_rows(output_table, distance_table)
  }
  return(output_table)
}

cell_cycle = c("G1", "earlyS", "lateS", "S_G2", "M", "asynch")

TPCA_Result = CorrAnalysis.TPCA(cell_cycle, becher, ori_et_al_complexes_df, filter_criteria, match = TRUE)
head(TPCA_Result, n=10)
```


**HMM Analysis**

#Initialize HMM model
Initialize the states, transitions, and emission matrix. 
```{r}
#Define HMM parameters

##States
states <- c("begin", "interaction", "no_interaction")

##Transitions
T_ <- matrix(c(0, 0, 0, 0.5, 0.75, 0.25, 0.5, 0.25, 0.75), nrow = 3) 
dimnames(T_) <- list(from = states, to = states) #names columns and rows for matrix
print(T_)
##Emissions
symbols = seq(from = 0.0, to = 1.0, by = 0.05)
probs = seq(from = 0.0, to = 1.0, by = 0.05)
probs_decrease_adjusted = probs * (1/sum(probs))
probs_increase_adjusted = rev(probs_decrease_adjusted)
E_ <- matrix(c(probs_increase_adjusted, probs_decrease_adjusted), 2, byrow=TRUE)
dimnames(E_) <- list(states = states[-1], residues = symbols) 
print(E_)

#creates hmm structure
Initial_HMM <- structure(list(A = T_, E = E_), class = "HMM")
plot.HMM(Initial_HMM, main ="Starting HMM Model for Protein Interactions")
```

#Run HMM analysis on the initial HMM
Run through all the proteins pairs in the complex of interest.
Use the Viterbi algorithm and HMM to determine true or false protien interactions.
```{r, warning=FALSE}
#Define the time points and order to perform HMM analysis
time_points = c("G1", "earlyS", "lateS", "S_G2", "M")

#Define dataframes
input_data = TPCA_Result
TPCA_HMM_Result = data.frame()

#Firt filter data by individual protien baits
baits = unique(input_data$bait)
for (i in baits) { 
  
  #Then filter by preys to run viterbi algorith on individual protein pairs
  input_data_bait = filter(input_data, bait == i)
  preys = unique(input_data_bait$prey)
  for (j in preys) { 
    
    #Filter and reshape data
    interaction_data =  filter(input_data_bait, prey == j) %>% filter(stage == time_points)
    observations = c(as.character(round(interaction_data$pvalue, 1)))
    
    #Run viterbi algorithm
    vit <- Viterbi(Initial_HMM, observations)
    predicted <- c(TRUE, FALSE)[vit$path +1]
    
    #Combine and create the full output
    interaction_data$interaction = predicted
    TPCA_HMM_Result = bind_rows(TPCA_HMM_Result, interaction_data)
  }
}

hist(TPCA_HMM_Result$distance, main = "Histogram of RSS input distances", xlab = "RSS")
barplot(table(TPCA_HMM_Result$interaction), main = "HMM Determined Protien Interaction")
head(TPCA_HMM_Result, n=10)

##Save the following file to then input into Cytoscape and create the protein networks
#write.csv(new_data, file = "exosome_hmm_ts75_tt25.csv", row.names = FALSE, quote = FALSE)
```


**Part 4: EM Optimization**

#EM Optimization: Aphid & RTPCA Data
Use the BaumWelch algorithm to train the emission and transition matrices. 
```{r}
#Create a training dataframe based on the observed data
observed_data <- c(as.character(round(TPCA_Result$distance, 1)))
observed_df <- data.frame(actual = c(TPCA_Result$distance))
observed_df$binned <- ifelse(observed_df$actual > 1, '1.0', observed_df$actual) 
train <- c(as.character(round(as.numeric(observed_df$binned, 1))))

#trains hmm -- use 1 set of observations not multiple observations
hmm_aphid <- train(Initial_HMM, train, method="BaumWelch", maxiter=500, pseudocounts="Laplace")

#Plot the emission matrices from before and after the Baum-Welch training
ggplot(data=data.frame(t(E_))) + 
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.05), y=interaction)) + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.05), y=interaction, color='Interaction')) +
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.05), y=no_interaction, )) + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.05), y=no_interaction, color='No Interaction')) +
  ggtitle("Initialized Emissions Matrix") + 
  labs(colour="Emission Matrix", x="Distances", y="Probability") +
  scale_color_manual(values = c('darkorange1', 'royalblue2')) +
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=data.frame(t(hmm_aphid$E))) + 
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.05), y=interaction)) + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.05), y=interaction, color='Interaction')) +
  geom_point(aes(x = seq(from = 0.0, to = 1.0, by = 0.05), y=no_interaction, )) + 
  geom_line(aes(x = seq(from = 0.0, to = 1.0, by = 0.05), y=no_interaction, color='No Interaction')) +
  ggtitle("Learned Emissions Matrix") + 
  labs(colour="Emission Matrix", x="Distances", y="Probability") +
  scale_color_manual(values = c('darkorange1', 'royalblue2')) +
  theme(legend.position="top") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#print out the table for the before and after of the transition matrices.
Initial_Transitions = data.frame(Initial_HMM$A)
print(Initial_Transitions)

EM_Learned_Transitions = data.frame(hmm_aphid$A)
print(EM_Learned_Transitions)

```

